---
title: "Functional Annotation with ClusterProfiler"
author: "Tania Quesada-López"
date: "03.03.2024 - 09.03.2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Overenrichment analysis.
Functional annotation of the DEGs with ClusterProfiler.

Automated label generation is allowed to associate these clusters with biological themes or processes of which the member gene sets are part of (such as `KEGG`, `Reactome`, and `Biocarta`pathways), based on a Jaccard index of a given threshold. 

1) Load data
The pathway analysis pipeline accepts count matrix files with """Ensembl ID""" as row
names. We can use our *Final Data* df.

2) Load necessary packages

```{r}
# To intall 
# devtools::install_github("YuLab-SMU/clusterProfiler")

require(clusterProfiler)
require(dbplyr)
require(sparklyr)
require(GSEA)
require(tibble)
library("RColorBrewer")
library("ggplot2")
library("gplots")
library("cowplot")
library("ggrepel")
library("reshape2")
library("knitr")
library("kableExtra")
library("igraph")
library("GGally")
library("DESeq2")
library("limma")
library("wordcloud")
library("tm")
library("SnowballC")
library("intergraph")
library("DT")
# library for mouse or human
# library("org.Hs.eg.db") # for human
library("org.Mm.eg.db") # for mouse

```

3) Run annotation 
Using log2FC and Symbol as rownames. GO annotation allows the use of *Symbol*. 

```{r}
# --- Over-representation analysis
organism <- org.Mm.eg.db

FinalData2 <- rownames_to_column(as.data.frame(FinalData)) # Get data with a column with the Symbol
df <- FinalData2[,c(1,26:31)] # Drop rows with info on expression
original_gene_list <- df$CE_vs_RTlog2FoldChange # Pull-out fold changes in response to cold
names(original_gene_list) <- df$rowname # Name those log2FoldChanges

gene_list = sort(original_gene_list, decreasing = TRUE) # ¡! Data should be sorted for ClusterProfiler

## Get GO terms for biological processes
gse <- gseGO(geneList=gene_list, 
             ont ="BP", 
             keyType = "SYMBOL", 
             minGSSize = 10, 
             maxGSSize = 1000, 
             pvalueCutoff = 0.05, 
             verbose = TRUE, 
             OrgDb = organism, 
             pAdjustMethod = "none")

foo <- as.data.frame(gse)

# write_xlsx(foo,  path =  "GO_RT.xlsx",
#           col_names = TRUE, format_headers = TRUE, use_zip64 = FALSE)

require(DOSE)
dotplot(gse, showCategory=10, split=".sign") + facet_grid(.~.sign)

ridgeplot(gse) + labs(x = "enrichment distribution")
# Use the `Gene Set` param for the index in the title, and as the value for geneSetId
gseaplot(gse, by = "all", title = gse$Description[2], geneSetID = 2)


gseBP <- gseGO(geneList=gene_list, 
               ont ="BP", 
               keyType = "SYMBOL", 
               nPerm = 10000, 
               minGSSize = 5, 
               maxGSSize = 800, 
               pvalueCutoff = 0.05, 
               verbose = TRUE, 
               OrgDb = organism, 
               pAdjustMethod = "BH")

foo <- summary(gseBP)
GO_BPordered <- arrange(foo, desc(enrichmentScore), p.adjust)
dotplot(gseBP, showCategory=10, split=".sign") + facet_grid(.~.sign)

```

4) Mapping Ensembl IDs to ENTREZ IDs
Maps *Symbol* to *ENTREZ IDs* wiht `org.Mm.eg.db` package (`UCSC` annotation database)
*Keep only genes with a unique mapping.
While GO annotation can worj with 


```{r}
# Convert gene IDs for gseKEGG function
# We will lose some genes here because not all IDs will be converted
colnames(sig_Condition)[1] <- c("SYMBOL") ## This df is ordered
data(geneList, package="DOSE")
keytypes(org.Mm.eg.db)
eg <- bitr(sig_Condition$SYMBOL, "SYMBOL", "ENTREZID", OrgDb = "org.Mm.eg.db")

gene_list2 <-eg[,2]

Res_Kegg <- enrichKEGG(gene_list2, organism = "mmu", keyType = "kegg", 
                       pvalueCutoff = 0.05, pAdjustMethod = "fdr", minGSSize = 10, maxGSSize = 500)

head(Res_Kegg)
foo <- summary(Res_Kegg)
KEGGordered <- arrange(foo, desc(qvalue))


dotplot(Res_Kegg)


```


